---
title: "Searching and Comparing PubMed Article Using R"
subtitle: "A simple practical user guides for publication statistics"
author: "<strong>Teresia Mrema Buza</strong>"
date: "Updated on `r Sys.Date()`"
site: bookdown::bookdown_site
knit: bookdown::render_book
output:
  bookdown::bs4_book:
    includes:
      in_header:
    template: bs4_template.html
documentclass: book
classoption: openany #remove empty pages in pdf doc
colorlinks: true
css:
- style.css
always_allow_html: true
fig_caption: true
fontsize: 12pt
geometry: margin=1in
indent: false
keep_tex: true
link-citations: true
mainfont: Times New Roman
biblio-style: apalike
description: |
  Tracking and comparing articles published in the NCBI PubMed Database.
---

# PubMed Publications Tracking {#pubmed-tracking}

```{r include=FALSE}
library(DiagrammeR)
library(DiagrammeRsvg)
library(rentrez)
library(glue)
library(tidyverse, suppressPackageStartupMessages)
library(tictoc)
library(xlsx)

## Default parameters
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  comment = NA,
  fig.path='./figures/',
  fig.show='asis',
  dev = 'png',
  fig.align='center',
  out.width = "70%",
  fig.width = 7,
  fig.asp = 0.7,
  fig.show = "asis"
)
```

## General Workflow
```{r echo=FALSE}
mermaidgraph <- mermaid("graph TD

subgraph  
A[<b>SEARCHING PUBMED ARTICLES</b>] -->  |Searching<br>Manually| B[Create Search Terms]
B --> C[NCBI PubMed Database] 
C --> |Results| C1[Indexed Papers in PubMed]
C1 --> |Download results| D[Results by Year in CSV format]
D --> |Filename| E[PubMed_Timeline_Results_by_Year.csv]
E --> F[Save the file in the data directory]
F -.-> G[Further Analysis in R<br>Such as plotting the results]


A --> |Seaching <br>Using a Code| H[Import Basic Libraries<li>rentrez</li><li>glue</li><li>tidyverse</li>]
H --> I[Define Search Terms]
I --> |Search counts| J[Computationally Get Search Counts]
J --> |Result<br> Manipulation| K[Tidying, Filtering, Transformation]
K --> L[Clean Data]
L --> F

end

", height = 800, width = 500)

mermaidgraph

```

## Structure of this guide {-}
This guide is divided into sections to facilitate easy navigation. The main chapter contains several sections as displayed in the navigation bars on the left and right. Click the hyper-linked text if you want to jump into a specific chapter or section.

## Code availability {-}
The code that generated this guide is available at a public [GitHub repository](https://github.com/tmbuza/pubmed-article-stats) or via a [GH-Page](https://tmbuza.github.io/pubmed-article-stats/). If interested you can request a consulting service by contacting the developer of this repo using <a href="https://complexdatainsights.com/contact-us">this contact form</a>.





